Point-process based representation learning for Electronic Health Records

### 2023-02-09 Manuscript_v1.tex
update

Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where each sample is represented as an event sequence  $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is number of event, $t_j$ is event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label). The history of events is denoted as $\mathcal{H}_t=\{(t_j,e_j):t_j<t \} $

@@

[ntpp] has proposed to extend for mult

marked cased

relation to autoencoder

@@

>>>>

\subsubsection{Temporal point process}
Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where each sample is represented as an event sequence  $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is number of event, $t_j$ is event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label). The history of events is denoted as $\mathcal{H}_t=\{(t_j,e_j):t_j<t \} $

@@

[ntpp] has proposed the following CIF to extend for multi-label case:

\begin{multline}\label{eq:11}
    -\log p(\mathcal{S}_{i})  =  - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log \lambda_m^{*}(t_j) \\   
    + \sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)\\
    - \sum_{j = 1}^{L}\sum_{m = 1}^{M} (1-1(e_j=m)) \log \left(1-\lambda_m^{*}(t_j) \exp \left( - \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right) \right)
\end{multline}

It should be noted that the real advantage of point process is modeling non-event likelihoods in the form of integrals. If we neglect the integrals, we would achieve the cross-entropy and binary cross entropy loss in the multi-class and multi-label settings respectively for the prediction of next mark given history of events.

\subsubsection{Neural temporal point process}

Encoder-decoder architectures have proven to be effective in many applications. The main idea of a neural point process is to first encode the history of events until $t_j$ using different neural network architectures $h_j=Enc(\mathcal{H}_j)$. Then it tries to estimate $\lambda^*_m(t)$ or its integral for $ t_i < t \leq t_{i+1}$.

\subsubsection{Deep learning for irregular sampling}
An irregularly sampled data can be denoted as $\mathcal{D}=\{\mathcal{U}_i\}_{i=1}^{N}$ where $N$ is number of samples and each sample is represented as series of tuple $\mathcal{U}_i={(t_p,k_p,v_p)}$ where $t_p, k_p, v_p$ represents the time, modality and value of $p$-th datapoint respectively.

@@

 The key advantage of our proposed model is to combine a  transformer-based event encoder with point process loss with a deep attention module that can handle an irregularly sampled time series for a any downstream prediciton task.
